Case-based Arabic Morphological Analyzer for Arabic Text Mining

### 2010-12-17 tex/sarf.aclhlt.tex
clarify alive, recursive, and controller early on

\begin{document}
\maketitle
\begin{abstract}
Natural language processing uses morphological 
analysis to abstract and annotate text.
Often times, 
%      and in particular in the context of the Arabic language, 
annotations resulting from 
a branch of morphological analysis may not be
appropriate for the case study at hand. 
In this paper we present Sarf, a case based morphological 
analyzer for Arabic. 
Sarf allows a case-based controller to control and refine
the state of the analysis on the fly. 
It uses non-deterministic finite state Automata
to analyze a stream of text. 
Each alive machine corresponds to a valid analysis. 
Sarf uses concatenative analysis based on recursive affixes 
to better preserve part of speech information.
We automated the analysis of three books of Islamic
narrations (hadith) using Sarf where we
abstracted each book into a vector of complex structures.
Our results show higher accuracy and better efficiency
compared to current analyzers.
\end{quote}
\end{abstract}

@@

{\bf Sarf.~~~}
In this paper, we present Sarf~\footnote{Sarf (name modified for blind review) 
will be available online as an open source tool.},
 a {\em novel case-based efficient
morphological analyzer} that uses 
non-deterministic finite state Automata 
driven by a case-based controller.
Each alive machine in Sarf represents a valid morphological analysis
and takes as input one letter at a time.
The case-based controller prunes false positives
early in the run by making a decision at every input letter
to control the alive machines.
%Each alive machine in Sarf takes as input one letter at a time 
%from a text stream. 
Sarf does not assume that the word at hand is a token and
performs tokenization on the fly based on morphological correctness.
Up to our knowledge, this is the first morphological analyzer that 
handles the ``run-on'' words problem.

Sarf uses concatenative analysis with a novel 
{\em recursive affixes} refinement. 
Sarf introduces recursive affixes in order to
better retain part of speech information and enhance the 
efficiency of affix matching.

@@

Similar to Sarf, Beesley~\shortcite{Beesley:01} uses
finite state Automata (FSA). 
It reports that the number of Automata generated by a compiler
for Xerox rules can not be controlled and also reports a difficulty 
to compose the FSAs into a single framework. 
Sarf constructs an NDFSA from
the controller and the prefix, stem, and suffix deterministic 
transducers to solve the problem.
Sarf differs in that it uses recursive affixes in the 
affix transducers, and in that the transducers consider
input from the controller to modify their behavior. 
%With Sarf it is easy to control the FSA with a case based
%controller. 
Beesley requires recompilation.% of the Xerox rules.

@@

Since $\Phi$ is now in an accept state in ${\cal S}$, it will
spawn $\Xi$. $\Xi$ transitions to the root of ${\cal X}$. 
$\Phi$ die when it consumes \RL{h} since
no edge from the current state is labeled with \RL{h}. 
$\Xi$ represents a valid analysis and can proceed.

@@

Atomic affixes are standalone affixes
that are not the concatenation of other affixes. 
Formally, we define an affix $f: a | a f$ to be either
an atomic affix $a$ or an atomic affix concatenated
with an affix. 
%Sarf only keeps atomic affixes and their categories.

@@

We investigated the case of Al-Kafi where the accuracy of 
the simple analysis was significantly lower than that of the 
refined analysis. 
We discovered that the chains detected by the refined analysis 
were a subset of those detected by the simple
analysis. 
The reason for the lower accuracy is false positive chains, 
implying that the simple analysis did not miss important 
information.

We hypothesize that a higher level of abstraction such as 
intersecting the narrators from the hadith books with the narrators 
in the biographies can reduce the false positives in the chains
in this case.

\subsection{Buckwalter and ElixirFM}

We implemented our case study with Buckwalter in two phases. 
In phase one, we processed the entire text with Buckwalter and
obtained annotations to text. 
In the second phase, we fed the annotations to 
the hadith controller as if they were coming from Sarf. 
We measured the time it took to accomplish both tasks
less the time spent streaming the annotations.
We did the same for ElixirFM.

Sarf outperforms both Buckwalter and ElixirFM in terms of 
running time. 
It also produces higher accuracy measures 
mainly due to the extended
lexicon as well as to the fact that Sarf reduces false positive
names by ignoring affixes that can not form
names even if one possible solution of the stem is a name. 
Doing the same with Buckwalter and ElixirFM requires  
expanding the annotations resulting from one simple affix to be 
reflected in all the composite affixes that contain it.
%expanding the annotations resulting from one affix to reflect
%its multiple composing affixes.
%However, this was easily performed using our recursive definition 
%of the affixes.

>>>>

\begin{document}
\maketitle
\begin{abstract}
Natural language processing uses morphological 
analysis to abstract and annotate text.
Often times, 
%      and in particular in the context of the Arabic language, 
annotations resulting from 
a branch of morphological analysis may not be
appropriate for the case study at hand. 
In this paper we present Sarf, a case-based morphological 
analyzer for Arabic. 
Sarf allows a case-based controller to control and refine
the state of the analysis on the fly. 
It uses non-deterministic finite state Automata
to analyze a stream of text. 
%Each machine corresponds to a valid analysis. 
Sarf uses concatenative analysis based on recursive affixes 
to better preserve part of speech information.
We automated the analysis of three books of Islamic
narrations (hadith) using Sarf where we
abstracted each book into a vector of complex structures.
Our results show higher accuracy and better efficiency
compared to current analyzers.
\end{quote}
\end{abstract}

@@

{\bf Sarf.~~~}
In this paper, we present Sarf~\footnote{Sarf (name modified for blind review) 
will be available online as an open source tool.},
 a {\em novel case-based efficient
morphological analyzer} that uses 
non-deterministic finite state Automata 
driven by a case-based controller.
Each machine in Sarf represents a valid morphological analysis
and takes as input one letter at a time.
The case-based controller prunes false positives
early in the run by making a decision at every input letter
to control the machines.
The controller we present in Section~\ref{sec:controller}
is an FSA. However, as we explain in Section~\ref{sec:ndfsa},
it can be a logical entity of any type.

%Each alive machine in Sarf takes as input one letter at a time 
%from a text stream. 
Sarf does not assume that the word at hand is a token and
performs tokenization on the fly based on morphological correctness.
Up to our knowledge, this is the first morphological analyzer that 
handles the ``run-on'' words problem.

Sarf uses concatenative analysis with a novel 
{\em recursive affixes} refinement. 
An atomic affix $a$ is a standalone affix
that is not the concatenation of other affixes. 
We define an affix $f: a | a f$ to be either
an atomic affix $a$ or an atomic affix concatenated
with an affix. 
Sarf introduces recursive affixes in order to
better retain part of speech information and enhance the 
efficiency of affix matching.

@@

Similar to Sarf, Beesley~\shortcite{Beesley:01} uses
finite state Automata (FSA). 
It reports that the number of Automata generated by a compiler
for Xerox rules can not be controlled and also reports a difficulty 
to compose the FSAs into a single framework. 
Sarf constructs an NDFSA from
the controller and the prefix, stem, and suffix deterministic 
transducers to solve the problem.
Sarf differs in that it uses recursive affixes in the 
affix transducers, and in that the transducers consider
input from the controller to modify their behavior. 
%With Sarf it is easy to control the FSA with a case-based
%controller. 
Beesley requires recompilation.% of the Xerox rules.

@@

Since $\Phi$ is now in an accept state in ${\cal S}$, it 
spawns $\Xi$ which transitions to the root of ${\cal X}$. 
After it consumes \RL{h--} $\Phi$ dies since there is no
\noTrRL{h--}-edge from the current state.
$\Xi$ represents a valid analysis and proceeds.

@@

%Sarf only keeps atomic affixes and their categories.

@@

We investigated the case of Al-Kafi where the accuracy of 
the simple analysis was significantly lower than that of the 
refined analysis. 
We discovered that the chains detected by the refined analysis 
were a subset of those detected by the simple
analysis. 
The reason for the lower accuracy is false positive chains, 
implying that the simple analysis did not miss important 
information. 
We hypothesize that a higher level of abstraction such as 
intersecting narrators from the hadith books with those from
the biographies can reduce the false positives in the chains.

%\subsection{Buckwalter and ElixirFM}
{\bf Buckwalter and ElixirFM.~~~}
We implemented our case study with Buckwalter in two phases. 
In phase one, we processed the entire text with Buckwalter and
obtained annotations to text. 
In the second phase, we fed the annotations to 
the hadith controller as if they were coming from Sarf. 
We measured the time it took to accomplish both tasks
less the time spent streaming the annotations.
We did the same for ElixirFM.

Sarf outperforms both Buckwalter and ElixirFM in terms of 
running time. 
It also produces higher accuracy measures 
mainly due to the extended
lexicon as well as to the fact that Sarf reduces false positive
names by ignoring affixes that can not form
names even if one possible solution of the stem is a name. 
%Doing the same with Buckwalter and ElixirFM requires  
%expanding the annotations resulting from one simple affix to be 
%reflected in all the composite affixes that contain it.
%expanding the annotations resulting from one affix to reflect
%its multiple composing affixes.
%However, this was easily performed using our recursive definition 
%of the affixes.
